Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-02-04 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Corrections

\subsection{Collaboration}
\label{sec:col}
Users need to authenticate themselves to the system using their
Facebook account. In that sense, some users might not be interested
either because they do not want to give that information or simply because
they do not use that social network. Even if we might lose
some users that way, the additional information we obtain for scoring
phenotypes more than compensates for that. 
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a rating to the phenotype or choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side of Figure \ref{fig:web}, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network. This makes the assignment 
of fitness through the rating system a {\em social} activity, 
pursuing the objective of this work, which is to increase engagement of the users.

@@

\subsection{Gamification}
The rewarding mechanism as it is applied in EvoDrawings gives more importance 
to the preference of those users with higher reputation
as given by their score points and experience levels.  
Each time a user does on of these actions their score is incremented by one:
\begin{itemize}
\item Start a session.
\item Rate a phenotype.
\item Create a collection.
\item Save a phenotype of the wall to a collection.
\item Save a phenotypes from a friend's collection.
\item Explore collections of other friends.
\end{itemize}

Two variables are used to determine the weight of a user's preference:
\begin{itemize}
\item {\em Experience}: This variable depends on the score and is a value 
between 0 and 100. It is assumed for this case, that once a user
reaches a 100 actions, it has enough experience on using the application.

\item {\em Participation}: This variable is simply the degree of the user node 
in the graph.    
\end{itemize}

@@

Three versions of EvoDrawings were compared:
\begin{itemize}
\item Base (B). All users have the same weight.
\item Non Graph Gamification (G). Only experience is considered
\item Graph Gamification (GG) . Both experience and participation are considered.
\end{itemize}
When gamification was employed, all score values known where presented to users
and a ranking of users by weight was shown in a window.
Table \ref{tab:params} shows the parameters used for the evolutionary algorithm.

@@

% Anuncements
At the start of every experiment, a call to participation was issued
through social networks.  The link used in the call for participation
was shortened Google URL, that provides metadata and analytics for the
users that click on it.  In table \ref{tab:urls} the URL for each 
deployment in the Heroku platform is shown,
along with the short URL and the analytics link. In the same table a link to the GitHub 
application repository used to deploy to Heroku is also listed.

Only data for the first week of deployment was considered 
for the experiments, and they where conducted 
between January and May of 2016.

\subsection{Results} 
Before launch, each deployment was first tried with a few beta testers. 
When applying the leader board gamification technique for the first time a 
problem was found in this stage. The problem was that some users were cheating by giving a
rating to an animation even before it was returned from the server, this was done by just
constantly clicking the mouse button. This is a common problem found in systems using leader
boards because by making the scores visible to other players they are encouraged 
to compete \cite{hickman2010total}. The results of each of the three experiments in 
terms of participation are detailed next.

@@

Table \ref{tab:urls} shows the total number of volunteers, nodes and edges 
in the graph after each experiment. Moreover, the total number of evaluated 
phenotypes for each volunteer is presented in figure 
\ref{fig:top-ranked-participation} where users are ranked by the 
number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
the number of phenotypes evaluated using a logarithmic scale. Results show that when 
considering deployments B and G, the difference came with users with a medium level of participation.
When comparing all the experiments the deployment GG had the higher number of participation in users of all levels.

@@

Moreover, modeling the performance of an C-IEC system
involves understanding its dynamics. Previous works on browser-based
volunteer computing have used metrics basic such as: 
the number of users or the time spent by every one in the
computation \cite{DBLP:journals/gpem/LaredoBGVAGF14, 2016arXiv160101607M}. 
While works on other platforms such as SETI@home
\cite{javadi2009mining} have found that the Weibull, log-normal, and
Gamma distributions where viable modeles of the  
the availability of computing resources in several clusters.
This is in concordance with the results obtained in \cite{agajaj}, and also
browser-based volunteer evolutionary systems like NodIO \cite{DBLP:conf/gecco/MereloCGCRV16}.
The shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas:
there are many users that give a small amount of cycles, while there
are just a few that give many cycles.

In orther to assess if C-IEC follow the same pattern, even when
using gamification techniques, participation was fittted to a 
Weibull distribution, results in figure \ref{fig:w2} and figure \ref{fig:w3}
show that the same pattern appears in this case study.

\begin{figure}[!t]
    \centering
        \includegraphics[width=3in]{img/weibull_2.png}
    \caption{Experiment 2. Weibull Fit}
    \label{fig:w2}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3in]{img/weibull_3.png}
    \caption{Experiment 3. Weibull Fit}
    \label{fig:w3}
\end{figure}

\section{Case Study: EvoDraw-Kinect}
\label{sec:kinect}
A variation of the previous application is The EvoDraw installation , using the same
framework but with a different form of interaction. Humans interact with the animations 
by simply looking at them. Again the population is stored in the EvoSpace container and
client computers remove animations from the EvoSpace to be presented in a display one at a time. 
After a certain amount of time,  animations are put back to the server along with information 
generated from the interaction. To continue, another is again removed and presented to viewers. 
An important feature is that once an animation is removed from the server, no other client can remove it, 
until the client that remove it puts the animation back, this means that any
animation is been shown only on one client at a given time. This cycle continues until there
is enough interactions to create a new generation of animations.

In order to assign a fitness value to each animation and thus enabling the EA the selection of the more interesting animations, a natural user interface device (Kinect sensor for Xbox One)
is used for facial tracking. Each animation is 
presented several times in a single generation and the best are always passed to the next, 
with its fitness recomputed. To generate the new population the same operators as the previous case study
are applied to the population.

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/kinect.png}
    \caption{EvoDraw Instalation on Alife XV Cancun Mexico.}
    \label{fig:kinect}
\end{figure}

In this instalation the components of the framework used where the EvoSpace and the Evolutionary Algorithm.
Information from users was limited to a stream of readings from the sensor.

@@

XYZ is a collective artwork generated by means of the Unplugged version of the Evolutionary Algorithm.
While in the previous cases software tools are in charge of running the EA and users collaborate evaluating
fitness quality, the sframework in XYZ is only in charge of storing digital versions of the works produced 
by the team of artists, who are really in charge of applying every step of the evolutionary algorithm. In 
this case only the relations are stored, and a basic user profile.

Thus, the main difference between the standard IEC and the Unplugged EA, is that in the latest,
every genetic operation as well as the general evolutionary loop is in charge of human beings. 
Thus the team of artists are in charge of:  (i) setting up the initial population by selecting
well known works from the history of arts;  (ii) each of the artists selects every generation two works 
from the previous one, and produce a new painting by applying any kind of crossover and mutation he may consider 
(iii) the process repeats for the number of generations initially decided, so that at the end, a collective 
artwork has been produced including as many paintings per generation, as artists are participating in the experiment.

Artists resort to framework as the tool allowing them to store the paintings, deciding their preferred ones from the
previous generation, as well as describing the reasons for their selections, mutation and crossover operations applied.
The tool is also in charge of generating statistics, computing \textit{elitism}, thus adding the most selected work to 
the next generation, etc.

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/interfaceXYZ.png}
    \caption{User interface of the XYZ application.}
    \label{fig:xyz}
\end{figure}

@@

Collaborative and volunteer based IEC systems involve the dynamic 
interaction of many entities and artifacts. Employing human centered approach 
can benefit researchers to understand and visualize this kind of systems better.
In this paper a human-centered software framework was proposed, and was validated
through the implementaion and refinement of three C-IEC applications.
This framework enabled the implementation of a gamification technique 
to improve engagement in a case study. Providing a common arena where users are aware of the
activities of other users in the social vicinity. Following a hybrid data model approach
gives developers and project owners options when innovating with other means
of interaction or new applications.

@@

\begin{acks}
  % The authors would like to thank Dr. Yuhua Li for providing the
  % matlab code of  the \textit{BEPS} method.

  % The authors would also like to thank the anonymous referees for
  % their valuable comments and helpful suggestions. The work is
  % supported by the \grantsponsor{GS501100001809}{National Natural
  %   Science Foundation of
  %   China}{http://dx.doi.org/10.13039/501100001809} under Grant
  % No.:~\grantnum{GS501100001809}{61273304}
  % and~\grantnum[http://www.nnsf.cn/youngscientsts]{GS501100001809}{Young
  % Scientsts' Support Program}.
  Reserved\\
  Space\\
  for\\
  Acks

>>>>

\subsection{Collaboration}
\label{sec:col}
Users need to authenticate themselves to the system using their
Facebook account. In that sense, some users might not be interested
either because they do not want to give that information or simply because
they do not use that social network. Even if we might lose
some users that way, the additional information we obtain for scoring
phenotypes more than compensates for that. 
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a rating to the phenotype or choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side of Figure \ref{fig:web}, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network. This makes the assignment 
of fitness through the rating system a {\em social} activity, 
pursuing the objective of this work, which is to increase user engagement.

@@

\subsection{Gamification}
The rewarding mechanism as it is applied in EvoDrawings gives more importance 
to the preference of those users with higher reputation
as given by their score points and experience levels.  
Each time a user does on of these actions their score is incremented by one:
start a session, rate a phenotype, create a collection, save a phenotype of 
the wall to a collection, save a phenotypes from a friend's collection, and
explore collections of other friends.

Two variables are used to determine the weight of a user's preference:
\begin{itemize}
\item {\bf Experience.}: This variable depends on the score and is a value 
between 0 and 100. It is assumed for this case, that once a user
reaches a 100 actions, it has enough experience on using the application.

\item {\bf Participation.}: This variable is simply the degree of the user node 
in the graph (number of edges).    
\end{itemize}

@@

Three versions of EvoDrawings were compared:
\begin{itemize}
\item Base (B): All users have the same weight.
\item Non Graph Gamification (G): Only experience is considered
\item Graph Gamification (GG): Both experience and participation are considered.
\end{itemize}
When gamification was employed, all score values known where presented to users
and a ranking of users by weight was shown in a window. Table \ref{tab:params} shows 
the parameters used for the evolutionary algorithm.

@@

% Anuncements
At the start of every experiment, a call to participation was issued
through social networks.  The link used in the call for participation
was shortened Google URL, that provides metadata and analytics for the
users that click on it.  In Table \ref{tab:urls} the URL for each 
deployment in the Heroku platform is shown,
along with the short URL and the analytics link. In the same table a link to the GitHub 
application repository used to deploy to Heroku is also listed.Only data 
for the first week of deployment was considered for the experiments, and they where conducted 
between January and May of 2016.

\subsection{Results} 
Before launch, each deployment was first tried with a few beta testers. 
When applying the leader board gamification technique for the first time a 
problem was found in this stage. The problem was that some users were cheating by giving a
rating to an animation even before it was returned from the server, this was done by just
constantly clicking the mouse button. This is a common problem found in systems using leader
boards because by making the scores visible to other players they are encouraged 
to compete \cite{hickman2010total}. The final version disabled the button until 
the drawing animation was over. The results of each of the three experiments in 
terms of participation are dtailed next.

@@

Table \ref{tab:urls} shows the total number of volunteers, nodes and edges 
in the graph after each experiment. Moreover, the total number of evaluated 
phenotypes for each volunteer is presented in figure 
\ref{fig:top-ranked-participation} where users are ranked by the 
number of phenotypes they rated. In the \emph{x} axis is the rank and in the \emph{y} axis 
the number of phenotypes evaluated using a logarithmic scale. Results show that when 
considering deployments B and G, the difference came with users with a medium level of participation.
When comparing all the experiments the deployment GG had the higher number of participation in users of all levels.

@@

Moreover, modeling the performance of an C-IEC system
involves understanding its dynamics. Previous works on browser-based
volunteer computing have used basic metrics such as: 
the number of users or the time spent by every one in the
computation \cite{DBLP:journals/gpem/LaredoBGVAGF14, 2016arXiv160101607M}. 
While works on other platforms such as SETI@home
\cite{javadi2009mining} have found that the Weibull, log-normal, and
Gamma distributions where viable modeles of the  
the availability of computing resources in several clusters.
This is in concordance with the results obtained in \cite{agajaj}, and also
browser-based volunteer evolutionary systems like NodIO \cite{DBLP:conf/gecco/MereloCGCRV16}.
The shape of those distributions is a skewed bell
with more resources in the {\em low} areas than in the high areas; i.e., 
there are many users that give a small amount of cycles, while there
are just a few that give many cycles.

In order to assess if C-IEC follows the same pattern, even when
using gamification techniques, participation was fittted to a 
Weibull distribution. Results in Figure \ref{fig:w2} and Figure \ref{fig:w3}
show that the same pattern appears in this case study.

\begin{figure}[!t]
    \centering
        \includegraphics[width=4in]{img/weibull_2.png}
    \caption{Experiment G. Weibull Fit}
    \label{fig:w2}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=4in]{img/weibull_3.png}
    \caption{Experiment GG. Weibull Fit}
    \label{fig:w3}
\end{figure}

\section{Case Study: EvoDraw-Kinect}
\label{sec:kinect}
A variation of the previous application is The EvoDraw installation, using the same
framework but with a different form of interaction. Humans interact with the animations 
by simply looking at them. Again the population is stored in the EvoSpace container and
client computers remove animations from the EvoSpace to be presented in a display one at a time. 
After a certain amount of time,  animations are put back to the server along with information 
generated from the interaction. To continue, another is again removed and presented to viewers. 
This cycle continues until there is enough interactions to create a new generation of animations.

In order to assign a fitness value to each animation and thus enabling 
the EA the selection of the most interesting animations, a natural user interface 
device (Kinect sensor for Xbox One) is used for facial tracking. Each animation is 
presented several times in a single generation and the best are always passed to the next, 
with its fitness recomputed. To generate the new population the same operators as the previous case study
are applied to the population.

\begin{figure}[!t]
    \centering
        \includegraphics[width=4in]{img/kinect.png}
    \caption{EvoDraw Instalation on Alife XV Cancun Mexico.}
    \label{fig:kinect}
\end{figure}

In this instalation the components of the framework used where EvoSpace and the EA.
Information from users was limited to a stream of readings from the sensor.

@@

XYZ is a collective artwork generated by means of the Unplugged version of the EA \cite{de2014artists}.
While in the previous cases interactive tools are in charge of running the EA and users collaborate evaluating
fitness quality, the proposed framework in XYZ is mainly in charge of storing digital versions of the works produced 
by the team of artists, who are really in charge of applying every step of the EA. Thus the team of artists 
are in charge of:  (i) setting up the initial population by selecting in the web interface
well known works from the history of arts;  (ii) each of the artists selects, every generation, two works 
from the previous one (see Figure \ref{fig:xyz}), and produce a new painting by applying 
any kind of crossover and mutation he or she may consider  (iii) the process repeats for the number 
of generations initially decided, so that at the end, a collective  artwork has been produced 
including as many paintings per generation, as artists are participating in the experiment. In this application
the final product is the motivation.

Artists resort to the framework as the tool allowing them to store the paintings, deciding their preferred ones from the
previous generation, as well as describing the reasons for their selections, mutation and crossover operations applied.
The tool is also in charge of generating statistics, computing \textit{elitism}, thus adding the most selected work to 
the next generation, etc.

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/interfaceXYZ.png}
    \caption{User interface of the XYZ application, used by artists to select the best pair of the current generation.}
    \label{fig:xyz}
\end{figure}

@@

Volunteer-based and C-IEC systems involve the dynamic 
interaction of many entities and artifacts. Employing a human-centered approach 
can benefit researchers to understand and visualize this kind of systems better.
In this paper a human-centered software framework was proposed, and was validated
through the implementaion and refinement of three C-IEC applications.
This framework enabled the implementation of a gamification technique 
to improve engagement in a case study. Providing a common arena where users are aware of the
activities of other users in the social vicinity. Following a hybrid data model approach
gives developers and project owners options when innovating with other means
of interaction or new applications.

@@

\begin{acks}
  Reserved\\
  Space\\
  for\\
  Acks
